Thesis: draft

### 2021-05-21 draft.tex
Update draft.tex
changed Ch.2 Watts-Strogatz model

Erd\H{o}s-R\'{e}nyi networks have small average distances and thus posses one of the two defining properties of the ``small world phenomenon''. Real social networks, however, usually display the ``small world phenomenon'' and thus have both short average distances and high clustering. The question then rises: Can we construct a model that has both of these properties? The answer is yes and this is what Duncan Watts and Steve Strogatz did in 1998. They constructed a network model that generates random networks with small world properties by interpolating between a regular ring lattice and a random network \cite{Watts1998}.\\
\newline
The construction of their model goes as follows \cite{Watts1998}: (i) Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side); (ii) For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $\beta$. Rewiring is done in such a way that self-loops and edge duplication are avoided.\\
\newline
The probability $\beta$ allows us to tune the network between a regular lattice ($\beta = 0$) and a random network ($\beta = 1$) \cite{Watts1998}. Usually, the following condition is required for the number of neighbors $K$: $N \gg K \gg \ln{N} \gg 1$. The first inequality assures sparse networks, whereas the second inequality guarantees that a random network will be connected (thus the network is not so sparse that it becomes disconnected) \cite{Watts1998}.\\
\newline
The regular ring lattice ($\beta=0$) has an average clustering coefficient $\left<cc(0)\right>$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
\begin{align}\label{C0}
	\left<cc(0)\right> = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 \ , && \left<d(0)\right> \approx \frac{N}{2K} \ .
\end{align}
The derivations of these formulas can be found in Appendix \ref{WS}, Section \ref{ringclus}. The regular ring lattice thus has a high clustering coefficient and an average path length that grows linearly with the system size $N$. \\
\newline
The random network ($\beta = 1$), on the other hand, has a low clustering coefficient $\left<cc(1)\right>$ and a small average path length $\left<d(1)\right>$ (Eq. (\ref{C}) and Eq. (\ref{d})) \cite{Watts1998}
\begin{align}
	\left<cc(1)\right> \approx \frac{K}{N} \ll 1 \ , && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}} \ .
\end{align}
Watts and Strogatz showed by numerical simulations that the average distance for the networks with $0 < \beta < 1$ is comparable with the average distance of a random network (which is small), even for small $\beta$ \cite{Newman2000}. The clustering coefficient is approximately given by
\begin{equation}
	\left<cc\right> \approx \left<cc(0)\right>(1-\beta)^3 \ ,
\end{equation} 
where $\left<cc(0)\right>$ is given by Eq. (\ref{C0}) \cite{Barrat1999}. This can be seen in the following way: (i) The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$; (ii) The probability that edges are rewired back to each other is negligible.\\
\newline
For small values of $\beta$, the clustering coefficient remains high. The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances. The Watts-Strogatz network has a homogeneous, bell shaped degree distribution with an average degree equal to $\left<k\right> = K$ \cite{Barrat1999}.

@@

\textbf{The Watts-Strogatz model}\\
\newline
The Watts-Strogatz model is used as a general model to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
\newline
Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}. Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn actors and we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient. We will investigate whether community structure behaves similar to high clustering regarding the appearance of echo chambers.\\
\newline
Beside the above, the Watts-Strogatz model is also used to build the constituent communities in the stochastic Watts-Strogatz block model.\\
\newline
\textbf{The stochastic block model}\\
\newline
One of the main features of the stochastic block model is the ability to generate networks with community structure. We will therefore use this model to investigate the impact of community structure on the evolution of opinions and the formation of echo chambers.\\
\newline
It is expected that community structure will enhance the formation of echo chambers compared to the random network case. The question remains, however, whether the influence of community structure on the formation of echo chambers is of the same caliber as that of high clustering. It is also of interest to study the possible relation between the appearance of echo chambers and the modularity of the network.\\
\newline
The stochastic Watts-Strogatz block model is used as a model that has both high clustering and community structure. We will check if the combination of both features displays any difference in the formation of echo chambers compared to having only a single feature (either community structure or high clustering).\\
\newline
All the above models have a homogeneous degree distribution. We did not focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model. It might, however, be an interesting extension to investigate such types of models as well.\\
\newline
The above models have their added value and benefits. One of their most important plus points is their simplicity. They are the most simple network models that one can create and each of them have characteristics and topologies that can be tuned (clustering for WS, community structure for SBM). This makes it really easy to investigate the effect of these different network topologies without the need to take the interplay with other possible complex network features and topologies into account. This plus side is, however, also one of the possible downsides. Real networks are not simple networks with single, isolated characteristics, but are instead complex networks where many different characteristics and topologies may each play a role. In other words, network models may lack many features of real-world networks. It is thus always tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world situations.\\
\newline
The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.

@@

ADD THIS IN AS LIMITATIONS/FUTURE DIRECTIONS

@@

>>>>

Erd\H{o}s-R\'{e}nyi networks have small average distances and thus posses one of the two defining properties of the ``small world phenomenon''. Real social networks, however, usually display the ``small world phenomenon'' and thus have both short average distances and high clustering. The question then rises whether we can construct a model that has both of these properties. The answer is yes and this is what Duncan Watts and Steve Strogatz did in 1998. They constructed a network model that generates random networks with small world properties by interpolating between a regular ring lattice and a random network \cite{Watts1998}.\\
The construction of their model goes as follows \cite{Watts1998}: (i) Start with a regular ring lattice with $N$ nodes, each connected to $K = 2m$ neighbors ($m$ on each side); (ii) For each node $i$, rewire each edge that connects $i$ with its $K/2$ rightmost neighbors with a probability $\beta$. Rewiring is done in such a way that self-loops and edge duplication are avoided.\\
The probability $\beta$ allows us to tune the network between a regular lattice ($\beta = 0$) and a random network ($\beta = 1$) \cite{Watts1998}. Usually, the following condition is required for the number of neighbors $K$: $N \gg K \gg \ln{N} \gg 1$. The first inequality assures sparse networks, whereas the second inequality guarantees that a random network will be connected (thus the network is not so sparse that it becomes disconnected) \cite{Watts1998}.\\
The regular ring lattice ($\beta=0$) has an average clustering coefficient $\left<cc(0)\right>$ and an average path length $\left<d(0)\right>$ equal to \cite{Watts1998}
\begin{align}\label{C0}
	\left<cc(0)\right> = \frac{3(K-2)}{4(K-1)} \approx \frac{3}{4} \ \text{for}\ K \gg 1 \ , && \left<d(0)\right> \approx \frac{N}{2K} \ .
The derivations of these formulas can be found in Appendix \ref{WS}, Section \ref{ringclus}. The regular ring lattice thus has a high clustering coefficient and an average path length that grows linearly with the system size $N$. \\
The random network ($\beta = 1$), on the other hand, has a low clustering coefficient $\left<cc(1)\right>$ and a small average path length $\left<d(1)\right>$ (Eq. (\ref{C}) and Eq. (\ref{d})) \cite{Watts1998}
	\left<cc(1)\right> \approx \frac{K}{N} \ll 1 \ , && \left<d(1)\right> \approx \frac{\ln{N}}{\ln{K}} \ .
Watts and Strogatz showed by numerical simulations that the average distance for the networks with $0 < \beta < 1$ is comparable with the average distance of a random network (which is small), even for small $\beta$ \cite{Newman2000}. The clustering coefficient is approximately given by
	\left<cc\right> \approx \left<cc(0)\right>(1-\beta)^3 \ ,
where $\left<cc(0)\right>$ is given by Eq. (\ref{C0}) \cite{Barrat1999}. This can be seen in the following way: (i) The probability that none of the three edges of the original triangles in the regular lattice are rewired is $(1-\beta)^3$; (ii) The probability that edges are rewired back to each other is negligible.\\
For small values of $\beta$, the clustering coefficient remains high. The model of Watts and Strogatz is thus able to generate random networks with a high clustering coefficient and short average distances. The Watts-Strogatz network has a homogeneous, bell shaped degree distribution with an average degree equal to $\left<k\right> = K$ \cite{Barrat1999}.\\
\textbf{The Watts-Strogatz model in this thesis}\\
The Watts-Strogatz model is used to investigate the effect of high clustering on the evolution of opinions and the formation of echo chambers.\\
Perra and Rocha showed that a high clustering coefficient enhances the formation of echo chambers \cite{Perra2019}. Here, we will use the Watts-Strogatz model to investigate the interplay between high clustering and the presence of stubborn actors and we will use it as a comparison for networks that posses a community structure without having a high clustering coefficient. We will investigate whether community structure behaves similar to high clustering regarding the appearance of echo chambers. The Watts-Strogatz model is also used to build the constituent communities in the stochastic Watts-Strogatz block model.

@@

\textbf{The stochastic block model}\\
\newline
One of the main features of the stochastic block model is the ability to generate networks with community structure. We will therefore use this model to investigate the impact of community structure on the evolution of opinions and the formation of echo chambers.\\
\newline
It is expected that community structure will enhance the formation of echo chambers compared to the random network case. The question remains, however, whether the influence of community structure on the formation of echo chambers is of the same caliber as that of high clustering. It is also of interest to study the possible relation between the appearance of echo chambers and the modularity of the network.\\
\newline
The stochastic Watts-Strogatz block model is used as a model that has both high clustering and community structure. We will check if the combination of both features displays any difference in the formation of echo chambers compared to having only a single feature (either community structure or high clustering).

@@

ADD THIS IN AS LIMITATIONS/FUTURE DIRECTIONS (about activation probabilities)

@@

+ ADD THIS (limitations of network models)

'''All the above models have a homogeneous degree distribution. We did not focus on network models which generate networks with a heterogeneous degree distribution such as the Barab\'{a}si-Albert model. It might, however, be an interesting extension to investigate such types of models as well.\\
\newline
The above models have their added value and benefits. One of their most important plus points is their simplicity. They are the most simple network models that one can create and each of them have characteristics and topologies that can be tuned (clustering for WS, community structure for SBM). This makes it really easy to investigate the effect of these different network topologies without the need to take the interplay with other possible complex network features and topologies into account. This plus side is, however, also one of the possible downsides. Real networks are not simple networks with single, isolated characteristics, but are instead complex networks where many different characteristics and topologies may each play a role. In other words, network models may lack many features of real-world networks. It is thus always tricky, and sometimes even incorrect, to extrapolate the results that are obtained by using network models to real-world situations.\\
\newline
The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.'''
